
```{r setup, include=FALSE}
setwd("<INSERT WORKING DIRECTORY>")
root = getwd()
knitr::opts_chunk$set(root = root)
knitr::opts_knit$set(root = root)
```

Libraries
```{r}
library(data.table)
library(stringr)
library(ggplot2)
library(dplyr)
```

First ext1f: Read in necessary files for ext1f.
```{r}
# .txt file generated from https://seer.cancer.gov/archive/csr/1975_2017/results_single/sect_01_table.04_2pgs.pdf
# seer data provided as txt
seer = fread("seer.txt", fill = T)

# lineage map provided as txt
lineage_match = fread("lineage_map.txt", header = T)
depmap_lineages = lineage_match[, 5:6]
seer_lineages = lineage_match[, 1:3]

ceres = fread(paste0(root, "/CRISPR_gene_effect.csv"))

sample_info = fread(paste0(root, "/sample_info.csv"))

#List of cell lines that also have RNA data to allow follow-up analysis using expression data if necessary
to_use = fread(paste0(root, "/samples_with_CERES_and_RNA.txt"))
```

Merge DepMap CDS2 or PARP dependency data with sample lineage info
```{r}
cds2 = grep("^CDS2",colnames(ceres))
cds2_ceres = ceres[, .(DepMap_ID, `CDS2 (8760)`)]
setnames(cds2_ceres, "CDS2 (8760)", "CDS2")
cds2_ceres = merge(cds2_ceres, sample_info[, .(DepMap_ID, CCLE_Name, primary_disease)], all.x = T)
```

Make a "lineage dictionary" to map all SEER labels to DepMap primary diseases
```{r}
lineage_dict = merge(depmap_lineages[!is.na(my_label)], seer_lineages[!is.na(my_label), .(my_label, site)])
setnames(lineage_dict, colnames(lineage_dict), c("new_label", "depmap_label", "seer_label"))
```

Make sure all lineage names overlap and include the brain mortality data under the appropriate label
```{r}
cds2_ceres[, new_label := lineage_dict[match(
  cds2_ceres[, str_replace(tolower(primary_disease), " ", "_")], depmap_label), 
  new_label] ]

seer_select = seer[seer$site %in% lineage_dict$seer_label, .(site, total_mortality, subcategory)]
seer_select[, new_label := lineage_dict[match(seer_select$site, seer_label), new_label]]

## brain does not have a total mortality
## replace with brain_nervous_system
seer_select[site=="Brain", total_mortality := 4.4]
```

Generate plot by merging DepMap and SEER data with CERES dependency data and plot information about CDS2 by calculating the percentage of CDS2-dependent cell lines by lineage.
```{r}
# collapse SEER by category (sum)
seer_select[, sum_tm := sum(as.numeric(total_mortality)), by = new_label]

#CDS2
depmap_seer = merge(cds2_ceres[!is.na(new_label), .(new_label, DepMap_ID, CDS2, primary_disease)], unique(seer_select[, .(new_label, sum_tm)]), by = "new_label")

depmap_seer[, lethality := ifelse(CDS2 < -0.5, "cds2.lethal", "cds2.nonlethal")]
depmap_seer = depmap_seer[DepMap_ID %in% to_use$RNA]
cell.lines = length(unique(depmap_seer$DepMap_ID))

plotting = dcast(depmap_seer[, .N, by = c("new_label", "lethality")], new_label ~ lethality)
plotting[is.na(cds2.lethal), cds2.lethal := 0]
plotting[, rate.by.lineage :=  log2(cds2.lethal + 1) ]
plotting[, rate.by.all :=  log2(((cds2.lethal/cell.lines))*100 + 1) ]

plotting = unique(merge(plotting, seer_select[, .(new_label, sum_tm)], by = "new_label"))
tm_all = as.numeric(seer[site=="All_Sites", total_mortality])
plotting[, mortality := log2(((as.numeric(sum_tm) / tm_all) * 100) + 1)]

plotting[, cds2_lethal_frac := cds2.lethal/(cds2.nonlethal + cds2.lethal)]
plotting[, cds2_lethal_frac_factor := ifelse(cds2_lethal_frac >= 0.5, ">=50%",
                                             ifelse(cds2_lethal_frac < 0.5 & cds2_lethal_frac >= 0.25, "25-50%",
                                                    ifelse(cds2_lethal_frac < 0.25 & cds2_lethal_frac >= 0.10, "10-25%", "<10%")))]

plotting[, cds2_lethal_frac_factor := factor(cds2_lethal_frac_factor, levels = c(">=50%", "25-50%", "10-25%", "<10%"))]
ggplot(data = plotting, aes(x = rate.by.all, y = mortality, col = cds2_lethal_frac_factor)) + geom_point(alpha = 0.2) + 
  geom_text(data = plotting, aes(x = rate.by.all, y = mortality, label = new_label))

plotted_data_CDS <- plotting
```

Read in necessary files for 1e and remove irrelevant files
```{r}
ccle = fread(paste0(root, "/CCLE_expression.csv"))
rm(cds2_ceres,ceres,depmap_lineages,lineage_dict,lineage_match,plotting,sample_info,seer,seer_lineages,seer_select,to_use,cds2,cell.lines,tm_all)
```

Collect expression data for 1e
```{r}
ccle_models <- ccle[,1]
CDS1_RNA <- ccle[ccle_models$V1 %in% depmap_seer$DepMap_ID,10883]
CDS1_RNA$DepMap_ID <- ccle_models$V1[ccle_models$V1 %in% depmap_seer$DepMap_ID]
```

Calculate expression % data for 1e
```{r}
ID_lineage <- depmap_seer[,1:2]

CDS1_RNA <- merge(CDS1_RNA,ID_lineage,by= "DepMap_ID")

CDS1_RNA[, poor_expression := ifelse(`CDS1 (1040)` < 1.5, "TRUE.rna", "FALSE.rna")]

Counts_RNA <- dcast(CDS1_RNA[, .N, by = c("new_label", "poor_expression")], new_label ~ poor_expression)

Counts_RNA[is.na(TRUE.rna), TRUE.rna := 0]

Counts_RNA[, rna_frac := TRUE.rna/(TRUE.rna + FALSE.rna)]
```

Collect data for 1e plot
```{r}
plot2_cds <- Counts_RNA[,1]
plot2_cds$rna_deficient_percentage <- data.frame(Counts_RNA$rna_frac*100)

lethal_frac_cds <- plotted_data_CDS[,c(1,8)]
lethal_frac_cds$cds2_lethal_percentage <- data.frame(lethal_frac_cds$cds2_lethal_frac*100)

#plot2_cds$new_label == lethal_frac_cds$new_label

plot2_cds$cds2_lethal_percentage <- lethal_frac_cds$cds2_lethal_percentage

rm(Counts_RNA,ID_lineage,lethal_frac_cds)
```

Now make plot 4d, ext4b

Remove irrelevant files
```{r}
rm(plotted_data_CDS)
```

Collect CDH1 and VIM expression data for
```{r}
CDH1_RNA <- ccle[ccle_models$V1 %in% depmap_seer$DepMap_ID,552]
CDH1_RNA$DepMap_ID <- ccle_models$V1[ccle_models$V1 %in% depmap_seer$DepMap_ID]

VIM_RNA <- ccle[ccle_models$V1 %in% depmap_seer$DepMap_ID,451]
VIM_RNA$DepMap_ID <- ccle_models$V1[ccle_models$V1 %in% depmap_seer$DepMap_ID]

ZEB1_RNA <- ccle[ccle_models$V1 %in% depmap_seer$DepMap_ID,8973]
ZEB1_RNA$DepMap_ID <- ccle_models$V1[ccle_models$V1 %in% depmap_seer$DepMap_ID]

rm(ccle,ccle_models)
```

Calculate expression % data
```{r}
ID_lineage <- depmap_seer[,1:2]

plot2_cds$rna_deficient_percentage <- NULL

CDS1_RNA[, poor_expression := ifelse(`CDS1 (1040)` > 2.5, "TRUE.rna", "FALSE.rna")]
Counts_RNA <- dcast(CDS1_RNA[, .N, by = c("new_label", "poor_expression")], new_label ~ poor_expression)
Counts_RNA[is.na(TRUE.rna), TRUE.rna := 0]
Counts_RNA[is.na(FALSE.rna), FALSE.rna := 0]
Counts_RNA[, rna_frac := TRUE.rna/(TRUE.rna + FALSE.rna)]
plot2_cds$CDS1_rna_proficient_percentage <- data.frame(Counts_RNA$rna_frac*100)
rm(Counts_RNA)

CDH1_RNA <- merge(CDH1_RNA,ID_lineage,by= "DepMap_ID")
CDH1_RNA[, poor_expression := ifelse(`CDH1 (999)` > 4.5, "TRUE.rna", "FALSE.rna")]
Counts_RNA <- dcast(CDH1_RNA[, .N, by = c("new_label", "poor_expression")], new_label ~ poor_expression)
Counts_RNA[is.na(TRUE.rna), TRUE.rna := 0]
Counts_RNA[is.na(FALSE.rna), FALSE.rna := 0]
Counts_RNA[, rna_frac := TRUE.rna/(TRUE.rna + FALSE.rna)]
plot2_cds$CDH1_rna_proficient_percentage <- data.frame(Counts_RNA$rna_frac*100)
rm(Counts_RNA)


VIM_RNA <- merge(VIM_RNA,ID_lineage,by= "DepMap_ID")
VIM_RNA[, poor_expression := ifelse(`VIM (7431)` > 6, "TRUE.rna", "FALSE.rna")]
Counts_RNA <- dcast(VIM_RNA[, .N, by = c("new_label", "poor_expression")], new_label ~ poor_expression)
Counts_RNA[is.na(TRUE.rna), TRUE.rna := 0]
Counts_RNA[is.na(FALSE.rna), FALSE.rna := 0]
Counts_RNA[, rna_frac := TRUE.rna/(TRUE.rna + FALSE.rna)]
plot2_cds$VIM_proficient_percentage <- data.frame(Counts_RNA$rna_frac*100)
rm(Counts_RNA)

ZEB1_RNA <- merge(ZEB1_RNA,ID_lineage,by= "DepMap_ID")
ZEB1_RNA[, poor_expression := ifelse(`ZEB1 (6935)` > 2, "TRUE.rna", "FALSE.rna")]
Counts_RNA <- dcast(ZEB1_RNA[, .N, by = c("new_label", "poor_expression")], new_label ~ poor_expression)
Counts_RNA[is.na(TRUE.rna), TRUE.rna := 0]
Counts_RNA[is.na(FALSE.rna), FALSE.rna := 0]
Counts_RNA[, rna_frac := TRUE.rna/(TRUE.rna + FALSE.rna)]
plot2_cds$ZEB1_proficient_percentage <- data.frame(Counts_RNA$rna_frac*100)
rm(Counts_RNA)

rm(CDH1_RNA, VIM_RNA, ZEB1_RNA, depmap_seer, ID_lineage)

write.csv(plot2_cds, "PlotExpand_CDS.csv")

```
