
```{r setup, include=FALSE}
setwd("<INSERT WORKING DIRECTORY>")
root = getwd()
knitr::opts_chunk$set(root = root)
knitr::opts_knit$set(root = root)
```

Load libraries
```{r}
library(data.table)
library(stringr)
library(plyr)
library(tidyr)
library(dplyr)
```

Load necessary files
```{r}
RNA_raw= read.csv(paste0(root, "/OmicsExpressionProteinCodingGenesTPMLogp1.csv"))
CHRONOS_raw = read.csv(paste0(root, "/CRISPRGeneEffect.csv"))
HGNC_raw = fread(paste0(root, "/hgnc_complete_set.txt"))
```

Format data
```{r}
RNA<-RNA_raw
CHRONOS<-CHRONOS_raw
rownames(RNA)<-RNA[,1]
rownames(CHRONOS)<-CHRONOS[,1]
RNA<-RNA[,-1]
CHRONOS<-CHRONOS[,-1]
non_na_counts <- colSums(!is.na(CHRONOS))
CHRONOS <- CHRONOS[, non_na_counts >= 500]
HGNC<-HGNC_raw[, .(entrez_id, symbol)]
PREVIOUS_HGNC <- HGNC_raw[,.(prev_symbol,symbol)]
rm(CHRONOS_raw,RNA_raw,HGNC_raw,non_na_counts)
```

Get complete cell line data (CHRONOS & RNA available)
```{r}
COMPLETE<-rownames(RNA)[rownames(RNA) %in% rownames(CHRONOS)]
RNA<-RNA[rownames(RNA) %in% COMPLETE,]
CHRONOS<-CHRONOS[rownames(CHRONOS) %in% COMPLETE,]
rm(COMPLETE)
```


Get entrez IDs
```{r}
new_colnames <- sub(".*\\.\\.", "", colnames(RNA)) # Remove ".." from column names
colnames(RNA) <- new_colnames
new_colnames <- sub(".*\\.\\.", "", colnames(CHRONOS)) # Remove ".." from column names
colnames(CHRONOS) <- new_colnames
colnames(RNA) <- str_remove(colnames(RNA), ".*\\.\\.")
colnames(CHRONOS) <- str_remove(colnames(CHRONOS), ".*\\.\\.")
colnames(RNA) <- sub("\\.$", "", colnames(RNA))
colnames(CHRONOS) <- sub("\\.$", "", colnames(CHRONOS))
rm(new_colnames)
```

Get the HGNC symbols for each entrez ID
```{r}
temp1 <- match(colnames(RNA), HGNC$entrez_id)
colnames(RNA)[!is.na(temp1)] <- as.character(HGNC$symbol[na.omit(temp1)])
temp2 <- match(colnames(CHRONOS), HGNC$entrez_id)
colnames(CHRONOS)[!is.na(temp2)] <- as.character(HGNC$symbol[na.omit(temp2)])
rm(temp1, temp2)
```

Combine transcripts for the same gene and remove transcripts that do not map to a gene (HGNC)
```{r}
RNA.T.names <- transpose(RNA, keep.names="colnames")
RNA.T <- RNA.T.names[,-1]
RNA.T.TPM <- (2**RNA.T)-1
RNA.T.TPM.aggregated <- aggregate(RNA.T.TPM, by = list(RNA.T.names$colnames), FUN = sum)
RNA.T.TPM.aggregated.genes <- RNA.T.TPM.aggregated[(RNA.T.TPM.aggregated$Group.1 %in% HGNC$symbol),]
RNA.T.aggregated.genes <- log2((RNA.T.TPM.aggregated.genes[,-1]+1))
RNA.aggregated.genes <- transpose(RNA.T.aggregated.genes)
colnames(RNA.aggregated.genes) <- RNA.T.TPM.aggregated$Group.1[RNA.T.TPM.aggregated$Group.1 %in% HGNC$symbol]
rownames(RNA.aggregated.genes) <- rownames(RNA)
RNA.aggregated.genes.orderedcol <- RNA.aggregated.genes[,order(colnames(RNA.aggregated.genes),decreasing=FALSE)]
RNA.aggregated.genes.orderedcolrow <- RNA.aggregated.genes.orderedcol[(order(rownames(RNA.aggregated.genes.orderedcol),decreasing=FALSE)),]
RNA<-RNA.aggregated.genes.orderedcolrow
rm(HGNC,RNA.T.TPM.aggregated,RNA.T.names,RNA.T,RNA.T.TPM,RNA.T.TPM.aggregated.genes,RNA.T.aggregated.genes, RNA.aggregated.genes,RNA.aggregated.genes.orderedcol,RNA.aggregated.genes.orderedcolrow)

write.table(RNA, paste0(root, "/RNA_aggregated_tpms_correct_genenames.txt"))
```

Calculate Pearson correlations and save intermediate results
```{r}
PEARSONS_pancancer <- cor(CHRONOS, RNA, method = c("pearson"), use = "pairwise.complete.obs")
write.table(PEARSONS_pancancer, file = paste0(root, "/PEARSONS_pancancer.txt"), sep = "\t")
rm(PEARSONS_pancancer)
```

Order results and remove duplicate anchor and target genes pancancer
```{r}
unmelted <- fread(paste0(root, "/PEARSONS_pancancer.txt"))
melted <- melt(unmelted, id.vars = "V1")
rm(unmelted)
melted.ordered <- melted[order(melted$value, decreasing=TRUE, )]
melted.ordered.V1 <- melted.ordered[!duplicated(melted.ordered$V1),]
results_pancancer <- melted.ordered.V1[!duplicated(melted.ordered.V1$variable),]
```

Write out results of the synthetic lethality prediction
```{r}
write.csv(results_pancancer, file = paste0(root, "/result_pancancer.csv"))
rm(results_pancancer)
```

calculate new significance cutt-off
```{r}
bonferroni_critical_correlation <- function(n_samples, n_tests) {
  # Bonferroni-corrected significance level
  alpha_corrected <- 0.05 / n_tests
  
  # Calculate the critical t-value for the corrected significance level
  critical_t <- qt(1 - alpha_corrected / 2, df = n_samples - 2)
  
  # Convert the critical t-value to a Pearson correlation coefficient
  critical_r <- critical_t / sqrt(critical_t^2 + (n_samples - 2))
  
  return(critical_r)
}

# Example usage:
n_samples <- 1021    # Number of samples
n_tests <- 17787*19153  # Number of tests to Bonferroni correct for

critical_r <- bonferroni_critical_correlation(n_samples, n_tests)
print(critical_r)
```
