
```{r setup, include=FALSE}
root = "<INSERT WORKING DIRECTORY>"
knitr::opts_chunk$set(root = root)
knitr::opts_knit$set(root = root)
```

Load libraries
```{r}
library(data.table)
library(stringr)
library(plyr)
```

Load necessary files
```{r}
RNA_raw= read.csv(paste0(root, "/CCLE_expression.csv"))
CERES_raw = read.csv(paste0(root, "/CRISPR_gene_effect.csv"))
HGNC_raw = fread(paste0(root, "/hgnc_complete_set.txt"))
INFO_raw = read.csv(paste0(root, "/sample_info.csv"))

TCGA_TUMOR_raw <- fread(paste0(root, "/GSM1536837_06_01_15_TCGA_24.tumor_Rsubread_TPM.txt"))
TCGA_HEALTHY_raw <- fread(paste0(root, "/GSM1697009_06_01_15_TCGA_24.normal_Rsubread_TPM.txt"))
```

Format data
```{r}
LINEAGE <- subset(INFO_raw, select =c("DepMap_ID","lineage"))
RNA<-RNA_raw
CERES<-CERES_raw
rownames(RNA)<-RNA[,1]
rownames(CERES)<-CERES[,1]
RNA<-RNA[,-1]
CERES<-CERES[,-1]
HGNC<-HGNC_raw[, .(entrez_id, symbol)]
PREVIOUS_HGNC <- HGNC_raw[,.(prev_symbol,symbol)]
rm(CERES_raw,INFO_raw, RNA_raw,HGNC_raw)
```

Get complete cell line data (CERES & RNA available)
```{r}
COMPLETE<-rownames(RNA)[rownames(RNA) %in% rownames(CERES)]
RNA<-RNA[rownames(RNA) %in% COMPLETE,]
CERES<-CERES[rownames(CERES) %in% COMPLETE,]
rm(COMPLETE)
```


Get entrez IDs
```{r}
new_colnames <- sub(".*\\.\\.", "", colnames(RNA)) # Remove ".." from column names
colnames(RNA) <- new_colnames
new_colnames <- sub(".*\\.\\.", "", colnames(CERES)) # Remove ".." from column names
colnames(CERES) <- new_colnames
colnames(RNA) <- str_remove(colnames(RNA), ".*\\.\\.")
colnames(CERES) <- str_remove(colnames(CERES), ".*\\.\\.")
colnames(RNA) <- sub("\\.$", "", colnames(RNA))
colnames(CERES) <- sub("\\.$", "", colnames(CERES))
rm(new_colnames)
```

Match cell lines and lineages
```{r}
lung <- LINEAGE$DepMap_ID[which(LINEAGE$lineage == "lung")]
skin <- LINEAGE$DepMap_ID[which(LINEAGE$lineage == "skin")]
breast <- LINEAGE$DepMap_ID[which(LINEAGE$lineage == "breast")]
cns <- LINEAGE$DepMap_ID[which(LINEAGE$lineage == "central_nervous_system")]
rm(LINEAGE)
```

Get the HGNC symbols for each entrez ID
```{r}
temp1 <- match(colnames(RNA), HGNC$entrez_id)
colnames(RNA)[!is.na(temp1)] <- as.character(HGNC$symbol[na.omit(temp1)])
temp2 <- match(colnames(CERES), HGNC$entrez_id)
colnames(CERES)[!is.na(temp2)] <- as.character(HGNC$symbol[na.omit(temp2)])
rm(temp1, temp2)
```

Combine transcripts for the same gene and remove transcripts that do not map to a gene (HGNC)
```{r}
RNA.T.names <- transpose(RNA, keep.names="colnames")
RNA.T <- RNA.T.names[,-1]
RNA.T.TPM <- (2**RNA.T)-1
RNA.T.TPM.aggregated <- aggregate(RNA.T.TPM, by = list(RNA.T.names$colnames), FUN = sum)
RNA.T.TPM.aggregated.genes <- RNA.T.TPM.aggregated[(RNA.T.TPM.aggregated$Group.1 %in% HGNC$symbol),]
RNA.T.aggregated.genes <- log2((RNA.T.TPM.aggregated.genes[,-1]+1))
RNA.aggregated.genes <- transpose(RNA.T.aggregated.genes)
colnames(RNA.aggregated.genes) <- RNA.T.TPM.aggregated$Group.1[RNA.T.TPM.aggregated$Group.1 %in% HGNC$symbol]
rownames(RNA.aggregated.genes) <- rownames(RNA)
RNA.aggregated.genes.orderedcol <- RNA.aggregated.genes[,order(colnames(RNA.aggregated.genes),decreasing=FALSE)]
RNA.aggregated.genes.orderedcolrow <- RNA.aggregated.genes.orderedcol[(order(rownames(RNA.aggregated.genes.orderedcol),decreasing=FALSE)),]
RNA<-RNA.aggregated.genes.orderedcolrow
rm(HGNC,RNA.T.TPM.aggregated,RNA.T.names,RNA.T,RNA.T.TPM,RNA.T.TPM.aggregated.genes,RNA.T.aggregated.genes, RNA.aggregated.genes,RNA.aggregated.genes.orderedcol,RNA.aggregated.genes.orderedcolrow)

write.table(RNA, paste0(root, "/RNA_aggregated_tpms_correct_genenames.txt"))
```

Calculate Pearson correlations and save intermediate results
```{r}
PEARSONS_pancancer <- cor(CERES, RNA, method = c("pearson"), use = "pairwise.complete.obs")
write.table(PEARSONS_pancancer, file = paste0(root, "/PEARSONS_pancancer.txt"), sep = "\t")
rm(PEARSONS_pancancer)

PEARSONS_lung <- cor(CERES[rownames(CERES) %in% lung,], RNA[rownames(RNA) %in% lung,], method = c("pearson"), use = "pairwise.complete.obs")
write.table(PEARSONS_lung, file = paste0(root, "/PEARSONS_lung.txt"), sep = "\t")
rm(PEARSONS_lung)

PEARSONS_skin <- cor(CERES[rownames(CERES) %in% skin,], RNA[rownames(RNA) %in% skin,], method = c("pearson"), use = "pairwise.complete.obs")
write.table(PEARSONS_skin, file = paste0(root, "/PEARSONS_skin.txt"), sep = "\t")
rm(PEARSONS_skin)

PEARSONS_breast <- cor(CERES[rownames(CERES) %in% breast,], RNA[rownames(RNA) %in% breast,], method = c("pearson"), use = "pairwise.complete.obs")
write.table(PEARSONS_breast, file = paste0(root,"/PEARSONS_breast.txt"), sep = "\t")
rm(PEARSONS_breast)

PEARSONS_cns <- cor(CERES[rownames(CERES) %in% cns,], RNA[rownames(RNA) %in% cns,], method = c("pearson"), use = "pairwise.complete.obs")
write.table(PEARSONS_cns, file = paste0(root, "/PEARSONS_cns.txt"), sep = "\t")
rm(PEARSONS_cns)

rm(breast, cns, lung, skin)
```

Order results and remove duplicate anchor and target genes pancancer
```{r}
unmelted <- fread(paste0(root, "/PEARSONS_pancancer.txt"))
melted <- melt(unmelted, id.vars = "V1")
rm(unmelted)
melted.ordered <- melted[order(melted$value, decreasing=TRUE, )]
melted.ordered.V1 <- melted.ordered[!duplicated(melted.ordered$V1),]
results_pancancer <- melted.ordered.V1[!duplicated(melted.ordered.V1$variable),]
```

Repeat for subtypes
```{r}
unmelted <- fread(paste0(root, "/PEARSONS_lung.txt"))
melted <- melt(unmelted, id.vars = "V1")
rm(unmelted)
melted.ordered <- melted[order(melted$value, decreasing=TRUE, )]
melted.ordered.V1 <- melted.ordered[!duplicated(melted.ordered$V1),]
results_lung <- melted.ordered.V1[!duplicated(melted.ordered.V1$variable),]

unmelted <- fread(paste0(root, "/PEARSONS_skin.txt"))
melted <- melt(unmelted, id.vars = "V1")
rm(unmelted)
melted.ordered <- melted[order(melted$value, decreasing=TRUE, )]
melted.ordered.V1 <- melted.ordered[!duplicated(melted.ordered$V1),]
results_skin <- melted.ordered.V1[!duplicated(melted.ordered.V1$variable),]

unmelted <- fread(paste0(root, "/PEARSONS_breast.txt"))
melted <- melt(unmelted, id.vars = "V1")
rm(unmelted)
melted.ordered <- melted[order(melted$value, decreasing=TRUE, )]
melted.ordered.V1 <- melted.ordered[!duplicated(melted.ordered$V1),]
results_breast <- melted.ordered.V1[!duplicated(melted.ordered.V1$variable),]

unmelted <- fread(paste0(root, "/PEARSONS_cns.txt"))
melted <- melt(unmelted, id.vars = "V1")
rm(unmelted)
melted.ordered <- melted[order(melted$value, decreasing=TRUE, )]
melted.ordered.V1 <- melted.ordered[!duplicated(melted.ordered$V1),]
results_cns <- melted.ordered.V1[!duplicated(melted.ordered.V1$variable),]
rm(melted, melted.ordered, melted.ordered.V1)
```

Write out results of the synthetic lethality prediction
```{r}
write.csv(results_pancancer, file = paste0(root, "/result_pancancer.csv"))
write.csv(results_lung, file = paste0(root, "/result_lung.csv"))
write.csv(results_skin, file = paste0(root, "/result_skin.csv"))
write.csv(results_breast, file = paste0(root, "/result_breast.csv"))
write.csv(results_cns, file = paste0(root, "/result_cns.csv"))
rm(results_pancancer,results_lung,results_skin,results_breast,results_cns, CERES, RNA)
```

Calculate TCGA fold changes

Format data
```{r}
TCGA_HEALTHY <- data.frame(TCGA_HEALTHY_raw)
rownames(TCGA_HEALTHY) <- TCGA_HEALTHY[,1]
TCGA_HEALTHY <- TCGA_HEALTHY[,-1]

TCGA_TUMOR <- data.frame(TCGA_TUMOR_raw)
rownames(TCGA_TUMOR) <- TCGA_TUMOR[,1]
TCGA_TUMOR <- TCGA_TUMOR[,-1]

rm(TCGA_HEALTHY_raw,TCGA_TUMOR_raw)
```


Calculate means
```{r}
healthy_means <- data.frame(apply(TCGA_HEALTHY,MARGIN=1,FUN=mean))
tumor_means <- data.frame(apply(TCGA_TUMOR,MARGIN=1,FUN=mean))
fold_changes <- tumor_means/healthy_means
```

Update HGNC symbols
```{r}
fold_changes.T <- transpose(fold_changes)
colnames(fold_changes.T) <- rownames(fold_changes)
colnames(fold_changes.T)[colnames(fold_changes.T) %in% PREVIOUS_HGNC$prev_symbol] <- na.omit(PREVIOUS_HGNC$symbol[match(colnames(fold_changes.T),PREVIOUS_HGNC$prev_symbol)])
fold_changes.updated <- transpose(fold_changes.T, keep.names="HGNC")
write.csv(fold_changes.updated, paste0(root, "/results_TCGA.csv"))
```

Genes of interest (goi) for panel ext1e
```{r}
genes_of_interest <- data.frame("CDS1","DDX3Y","EIF1AY","FAM50B")
healthygoi <- TCGA_HEALTHY[rownames(TCGA_HEALTHY) %in% genes_of_interest,]
tumorgoi <- TCGA_TUMOR[rownames(TCGA_TUMOR) %in% genes_of_interest,]
write.csv(healthygoi, paste0(root, "/healthygoi.csv"))
write.csv(tumorgoi, paste0(root, "/tumorgoi.csv"))
```