
```{r setup, include=FALSE}
setwd("<INSERT WORKING DIRECTORY>")
root = getwd()
knitr::opts_chunk$set(root = root)
knitr::opts_knit$set(root = root)
```

Load libraries
```{r}
library(data.table)
library(stringr)
library(plyr)
```

Load files
```{r}
CHRONOS_raw = fread(paste0(root, "/CRISPRGeneEffect.csv"))
RNA_raw = fread(paste0(root, "/OmicsExpressionProteinCodingGenesTPMLogp1.csv"))
MUT_DAM_raw = fread(paste0(root, "/OmicsSomaticMutationsMatrixDamaging.csv"))
MUT_HOT_raw = fread(paste0(root, "/OmicsSomaticMutationsMatrixHotspot.csv"))
HGNC_raw = fread(paste0(root, "/hgnc_complete_set.txt"))
```

Format data
```{r}
RNA<-data.frame(RNA_raw)
rownames(RNA)<-RNA[,1]
RNA<-RNA[,-1]

MUT_HOT<-data.frame(MUT_HOT_raw)
rownames(MUT_HOT)<-MUT_HOT[,1]
MUT_HOT<-MUT_HOT[,-1]

MUT_DAM<-data.frame(MUT_DAM_raw)
rownames(MUT_DAM)<-MUT_DAM[,1]
MUT_DAM<-MUT_DAM[,-1]

CHRONOS<-data.frame(CHRONOS_raw)
rownames(CHRONOS)<-CHRONOS[,1]
CHRONOS<-CHRONOS[,-1]
non_na_counts <- colSums(!is.na(CHRONOS))
CHRONOS <- CHRONOS[, non_na_counts >= 500]

HGNC<-HGNC_raw[, .(entrez_id, symbol)]
PREVIOUS_HGNC <- HGNC_raw[,.(prev_symbol,symbol)]
rm(MUT_HOT_raw,MUT_DAM_raw, RNA_raw,HGNC_raw,CHRONOS_raw,non_na_counts)
```

Get complete cell line data (CHRONOS & RNA available)
```{r}
COMPLETE<-rownames(RNA)[rownames(RNA) %in% rownames(CHRONOS)]
RNA<-RNA[rownames(RNA) %in% COMPLETE,]
CHRONOS<-CHRONOS[rownames(CHRONOS) %in% COMPLETE,]
MUT_DAM<-MUT_DAM[rownames(MUT_DAM) %in% COMPLETE,]
MUT_HOT<-MUT_HOT[rownames(MUT_HOT) %in% COMPLETE,]
rm(COMPLETE)
```

Get entrez IDs
```{r}
new_colnames <- sub(".*\\.\\.", "", colnames(RNA)) # Remove ".." from column names
colnames(RNA) <- new_colnames
colnames(RNA) <- str_remove(colnames(RNA), ".*\\.\\.")
colnames(RNA) <- sub("\\.$", "", colnames(RNA))

new_colnames <- sub(".*\\.\\.", "", colnames(CHRONOS)) # Remove ".." from column names
colnames(CHRONOS) <- new_colnames
colnames(CHRONOS) <- str_remove(colnames(CHRONOS), ".*\\.\\.")
colnames(CHRONOS) <- sub("\\.$", "", colnames(CHRONOS))

new_colnames <- sub(".*\\.\\.", "", colnames(MUT_DAM)) # Remove ".." from column names
colnames(MUT_DAM) <- new_colnames
colnames(MUT_DAM) <- str_remove(colnames(MUT_DAM), ".*\\.\\.")
colnames(MUT_DAM) <- sub("\\.$", "", colnames(MUT_DAM))

new_colnames <- sub(".*\\.\\.", "", colnames(MUT_HOT)) # Remove ".." from column names
colnames(MUT_HOT) <- new_colnames
colnames(MUT_HOT) <- str_remove(colnames(MUT_HOT), ".*\\.\\.")
colnames(MUT_HOT) <- sub("\\.$", "", colnames(MUT_HOT))

rm(new_colnames)
```

Get the HGNC symbols for each entrez ID
```{r}
temp <- match(colnames(RNA), HGNC$entrez_id)
colnames(RNA)[!is.na(temp)] <- as.character(HGNC$symbol[na.omit(temp)])

temp <- match(colnames(CHRONOS), HGNC$entrez_id)
colnames(CHRONOS)[!is.na(temp)] <- as.character(HGNC$symbol[na.omit(temp)])

temp <- match(colnames(MUT_DAM), HGNC$entrez_id)
colnames(MUT_DAM)[!is.na(temp)] <- as.character(HGNC$symbol[na.omit(temp)])

temp <- match(colnames(MUT_HOT), HGNC$entrez_id)
colnames(MUT_HOT)[!is.na(temp)] <- as.character(HGNC$symbol[na.omit(temp)])

rm(temp)
```

Combine transcripts for the same gene and remove transcripts that do not map to a gene (HGNC)
```{r}
RNA.T.names <- transpose(RNA, keep.names="colnames")
RNA.T <- RNA.T.names[,-1]
RNA.T.TPM <- (2**RNA.T)-1
RNA.T.TPM.aggregated <- aggregate(RNA.T.TPM, by = list(RNA.T.names$colnames), FUN = sum)
RNA.T.TPM.aggregated.genes <- RNA.T.TPM.aggregated[(RNA.T.TPM.aggregated$Group.1 %in% HGNC$symbol),]
RNA.T.aggregated.genes <- log2((RNA.T.TPM.aggregated.genes[,-1]+1))
RNA.aggregated.genes <- transpose(RNA.T.aggregated.genes)
colnames(RNA.aggregated.genes) <- RNA.T.TPM.aggregated$Group.1[RNA.T.TPM.aggregated$Group.1 %in% HGNC$symbol]
rownames(RNA.aggregated.genes) <- rownames(RNA)
RNA.aggregated.genes.orderedcol <- RNA.aggregated.genes[,order(colnames(RNA.aggregated.genes),decreasing=FALSE)]
RNA.aggregated.genes.orderedcolrow <- RNA.aggregated.genes.orderedcol[(order(rownames(RNA.aggregated.genes.orderedcol),decreasing=FALSE)),]
RNA<-RNA.aggregated.genes.orderedcolrow
rm(HGNC,RNA.T.TPM.aggregated,RNA.T.names,RNA.T,RNA.T.TPM,RNA.T.TPM.aggregated.genes,RNA.T.aggregated.genes, RNA.aggregated.genes,RNA.aggregated.genes.orderedcol,RNA.aggregated.genes.orderedcolrow)

write.table(RNA, paste0(root, "/RNA_aggregated_tpms_correct_genenames.txt"))
```

Remove data for genetic sequences not (yet) officially recognized as genes
```{r}
MUT_DAM <- MUT_DAM[, !colnames(MUT_DAM) %in% "Unknown"]
MUT_HOT <- MUT_HOT[, !colnames(MUT_HOT) %in% "Unknown"]
```

Order data
```{r}
CHRONOS <- CHRONOS[order(rownames(CHRONOS)), ]
RNA <- RNA[order(rownames(RNA)), ]
MUT_DAM <- MUT_DAM[order(rownames(MUT_DAM)), ]
MUT_HOT <- MUT_HOT[order(rownames(MUT_HOT)), ]

#identical(rownames(RNA), rownames(CHRONOS)) && identical(rownames(MUT_DAM), rownames(MUT_HOT)) && identical(rownames(MUT_DAM), rownames(RNA))
```

Prune and select data
```{r}
CDS2_CHRONOS <- CHRONOS[colnames(CHRONOS)=="CDS2"]
CDS1_RNA <- RNA[colnames(RNA)=="CDH1"]
rm(CHRONOS, PREVIOUS_HGNC)
```

Calculate partial correlations expression
```{r}
CDS2chron_CDS1rna.XZ <- cor(CDS2_CHRONOS, CDS1_RNA, method = c("pearson"), use = "pairwise.complete.obs")

#Save for plotting later
ReferencePearson <- CDS2chron_CDS1rna.XZ

GENErna_CDS1rna.YZ <- cor(RNA, CDS1_RNA, method = c("pearson"), use = "pairwise.complete.obs")
CDS2chron_GENErna.XY <- cor(CDS2_CHRONOS, RNA, method = c("pearson"), use = "pairwise.complete.obs")
CDS2chron_GENErna.XY <- t(CDS2chron_GENErna.XY)

YZ <- GENErna_CDS1rna.YZ[, 1]
XY <- CDS2chron_GENErna.XY[, 1]

YZ <- YZ[!is.na(YZ)]
XY <- XY[!is.na(XY)]
identical(names(YZ),names(XY))

XZ <- CDS2chron_CDS1rna.XZ[1, 1]
rm(CDS2chron_CDS1rna.XZ)
XZ <- rep(XZ, 19151)
XZ <- c(XZ)
names(XZ) <- names(YZ)

# partial correlation between XY corrected for Z: XY.Z = (XY-(XZ*YZ))/(SQRT(1-YZ^2)*SQRT(1-XZ^2)), where two letter pairs indicate pearson correlations between the two variables and a dot followed by a symbol indicates the variable for which the correlation was corrected. (SQRT refers to taking square root, * is multiplication)

XY.Z <- (XY - (XZ * YZ)) / (sqrt(1 - (YZ^2)) * sqrt(1 - (XZ^2)))
Partial_Correlations <- data.frame(XY.Z)
colnames(Partial_Correlations) <- "Expression"

rm(XY, XY.Z, XZ, YZ, CDS2chron_GENErna.XY,GENErna_CDS1rna.YZ,RNA)
```

Calculate partial correlations damaging mutations
```{r}
MUT_DAM <- MUT_DAM[, colSums(MUT_DAM) > 0]

CDS2chron_CDS1rna.XZ <- cor(CDS2_CHRONOS, CDS1_RNA, method = c("pearson"), use = "pairwise.complete.obs")
GENEmutdam_CDS1rna.YZ <- cor(MUT_DAM, CDS1_RNA, method = c("pearson"), use = "pairwise.complete.obs")
CDS2chron_GENEmutdam.XY <- cor(CDS2_CHRONOS, MUT_DAM, method = c("pearson"), use = "pairwise.complete.obs")
CDS2chron_GENEmutdam.XY <- t(CDS2chron_GENEmutdam.XY)

YZ <- GENEmutdam_CDS1rna.YZ[, 1]
XY <- CDS2chron_GENEmutdam.XY[, 1]

YZ <- YZ[!is.na(YZ)]
XY <- XY[!is.na(XY)]
identical(names(YZ),names(XY))

XZ <- CDS2chron_CDS1rna.XZ[1, 1]
rm(CDS2chron_CDS1rna.XZ)
XZ <- rep(XZ, 15421)
XZ <- c(XZ)
names(XZ) <- names(YZ)

# partial correlation between XY corrected for Z: XY.Z = (XY-(XZ*YZ))/(SQRT(1-YZ^2)*SQRT(1-XZ^2)), where two letter pairs indicate pearson correlations between the two variables and a dot followed by a symbol indicates the variable for which the correlation was corrected. (SQRT refers to taking square root, * is multiplication)

XY.Z <- (XY - (XZ * YZ)) / (sqrt(1 - (YZ^2)) * sqrt(1 - (XZ^2)))
Partial_Correlations_2 <- data.frame(XY.Z)
colnames(Partial_Correlations_2) <- "MUT_DAM"

rm(XY, XY.Z, XZ, YZ, CDS2chron_GENEmutdam.XY,GENEmutdam_CDS1rna.YZ,MUT_DAM)
```

Calculate partial correlations hotspot mutations
```{r}
MUT_HOT <- MUT_HOT[, colSums(MUT_HOT) > 0]

CDS2chron_CDS1rna.XZ <- cor(CDS2_CHRONOS, CDS1_RNA, method = c("pearson"), use = "pairwise.complete.obs")
GENEmuthot_CDS1rna.YZ <- cor(MUT_HOT, CDS1_RNA, method = c("pearson"), use = "pairwise.complete.obs")
CDS2chron_GENEmuthot.XY <- cor(CDS2_CHRONOS, MUT_HOT, method = c("pearson"), use = "pairwise.complete.obs")
CDS2chron_GENEmuthot.XY <- t(CDS2chron_GENEmuthot.XY)

YZ <- GENEmuthot_CDS1rna.YZ[, 1]
XY <- CDS2chron_GENEmuthot.XY[, 1]

YZ <- YZ[!is.na(YZ)]
XY <- XY[!is.na(XY)]
identical(names(YZ),names(XY))

XZ <- CDS2chron_CDS1rna.XZ[1, 1]
rm(CDS2chron_CDS1rna.XZ)
XZ <- rep(XZ, 372)
XZ <- c(XZ)
names(XZ) <- names(YZ)

# partial correlation between XY corrected for Z: XY.Z = (XY-(XZ*YZ))/(SQRT(1-YZ^2)*SQRT(1-XZ^2)), where two letter pairs indicate pearson correlations between the two variables and a dot followed by a symbol indicates the variable for which the correlation was corrected. (SQRT refers to taking square root, * is multiplication)

XY.Z <- (XY - (XZ * YZ)) / (sqrt(1 - (YZ^2)) * sqrt(1 - (XZ^2)))
Partial_Correlations_3 <- data.frame(XY.Z)
colnames(Partial_Correlations_3) <- "MUT_HOT"

rm(XY, XY.Z, XZ, YZ, CDS2chron_GENEmuthot.XY,GENEmuthot_CDS1rna.YZ,MUT_HOT,CDS1_RNA,CDS2_CHRONOS)
```

Merge data
```{r}
Partial_Correlations["Gene"] <- rownames(Partial_Correlations)
rownames(Partial_Correlations) <- NULL
EXP <- melt(Partial_Correlations, variable.name = "Data Type", value.name = "Partial Correlations")

Partial_Correlations_2["Gene"] <- rownames(Partial_Correlations_2)
rownames(Partial_Correlations_2) <- NULL
MUT_DAM <- melt(Partial_Correlations_2, variable.name = "Data Type", value.name = "Partial Correlations")

Partial_Correlations_3["Gene"] <- rownames(Partial_Correlations_3)
rownames(Partial_Correlations_3) <- NULL
MUT_HOT <- melt(Partial_Correlations_3, variable.name = "Data Type", value.name = "Partial Correlations")

Results <- rbind(EXP, MUT_DAM, MUT_HOT)
Results <- Results[!is.na(Results$`Partial Correlations`),]
rm(EXP,MUT_DAM,MUT_HOT,Partial_Correlations,Partial_Correlations_2,Partial_Correlations_3)
```

Calculate significance
```{r}
pearson_p_value <- function(correlation, n) {
  t_stat <- correlation * sqrt((n - 2) / (1 - correlation^2))
  p_value <- 2 * pt(-abs(t_stat), df = n - 2)
  return(p_value)
}
Results["P-values_corrected"] <- pmin(pearson_p_value(Results$`Partial Correlations`,1021)*34943,1)
Results <- Results[order(Results$`P-values_corrected`), ]
Results <- Results[order(-Results$`Partial Correlations`), ]
write.csv(Results,"Results.csv")

#Significance cut-off Pearson in plot

pearson_from_p_value <- function(p_value, n) {
  t_value <- qt(p_value / 2, df = n - 2, lower.tail = FALSE)
  r_value <- t_value / sqrt(t_value^2 + (n - 2))
  return(r_value)
}

pearson_from_p_value(0.05/34943,1021)
# cut-offs are 0.1501939 and -0.1501939

#Strength of CDS1 RNA CDS2 CHRONOS Pearson for reference
ReferencePearson
#0.69 is the reference

rm(ReferencePearson, Results,pearson_p_value,pearson_from_p_value)
```

```{r}
sessionInfo()
```

