
```{r setup, include=FALSE}
root = "<INSERT WORKING DIRECTORY>"
knitr::opts_chunk$set(root = root)
knitr::opts_knit$set(root = root)
```

Load libraries
```{r}
library(data.table)
library(stringr)
library(ggplot2)
library(cowplot)
```

Read in necessary files
```{r}
RNA = fread(paste0(root, "/RNA_aggregated_tpms_correct_genenames.txt")) # we generate this in the code for panel 1c
tcga.expr = fread(paste0(root, "/EBPlusPlusAdjustPANCAN_IlluminaHiSeq_RNASeqV2.geneExp.tsv"))
HGNC = fread(paste0(root, "/hgnc_complete_set.txt"))
CERES = read.csv(paste0(root, "/CRISPR_gene_effect.csv"))
```

Prepare TCGA data with correct gene names
```{r}
tcga.expr[, entrez_id := str_extract(gene_id, "(?<=|)[0-9]+$")]

tcga.expr[, hgnc_new := HGNC[match(tcga.expr$entrez_id, entrez_id), symbol]]
tcga.expr = tcga.expr[!is.na(hgnc_new)]
```

Pearson correlation between CDS1 expression and expression every other gene DepMap
```{r}
RNA = as.data.frame(RNA)
rownames(RNA) = RNA$V1
RNA$V1 = NULL
cds1_pos = which(colnames(RNA)=="CDS1")
gene_cor = as.data.frame(t(cor(RNA[, cds1_pos], RNA, use = "pairwise.complete.obs")))
```

Pearson between CDS1 expression and expression of every other gene TCGA
```{r}
tcga.expr[, `:=` (entrez_id=NULL,
                   gene_id=NULL)]
tcga = as.data.frame(tcga.expr)
rownames(tcga) = tcga$hgnc_new
tcga$hgnc_new = NULL

tcga = t(tcga)

cds1_pos_tcga = which(colnames(tcga)=="CDS1")
gene_cor_tcga = as.data.frame(t(cor(tcga[, cds1_pos_tcga], tcga, use = "pairwise.complete.obs")))
```

Pearson between CDS2 dependency and expression of every other gene DepMap
```{r}
CERES = as.data.frame(CERES)
rownames(CERES) = CERES$DepMap_ID
CERES$DepMap_ID = NULL

cds2_pos = which(grepl("CDS2", colnames(CERES)))
cds2_dep = CERES[,cds2_pos, drop=F]
colnames(cds2_dep) = "CDS2"

cds2 = cbind(cds2_dep, RNA[match(rownames(cds2_dep), rownames(RNA)),])

cor_dep = t(as.data.frame(cor(cds2[, 1], cds2[,-1], use = "pairwise.complete.obs")))
```

Plot
```{r}
colnames(cor_dep) = "cds2_dep"
colnames(gene_cor) = "rna"
plot = cbind(as.data.frame(gene_cor), as.data.frame(cor_dep))
plot$label = rownames(plot)

ggplot(data = plot, aes(x = rna, y = cds2_dep)) + geom_point(alpha = 0.1) + xlim(-1,1) + ylim(-1,1) + theme_cowplot() + geom_point(data = plot[rownames(plot) %in% c("VIM", "ZEB1", "ZEB2", "CDS1", "CDH1"),], col = "red") + ggrepel::geom_label_repel(data = plot[rownames(plot) %in% c("VIM", "ZEB1", "ZEB2", "CDS1", "CDH1"),], aes(label = label))
```

Write out .rnk files for GSEA on DepMap and TCGA CDS1 co-expression Pearson correlations
```{r}
fwrite(gene_cor, col.names = F, quote = F, sep = "\t", file = paste0(root, "/depmap.rnk"), row.names = T)
fwrite(gene_cor_tcga, col.names = F, quote = F, sep = "\t", file = paste0(root, "/tcga.rnk"), row.names = T)
```
