
```{r setup, include=FALSE}
root = "<INSERT WORKING DIRECTORY>"
knitr::opts_chunk$set(root = root)
knitr::opts_knit$set(root = root)
```

Load libraries
```{r}
library(data.table)
```

RNA tpms with correct names
```{r}
RNA = fread(paste0(root, "/RNA_aggregated_tpms_correct_genenames.txt")) # we generate this in the code for panel 1c
HGNC_raw = fread(paste0(root, "/hgnc_complete_set.txt"))
CERES_raw <- read.csv(paste0(root, '/CRISPR_gene_effect.csv'))
drugs = read.delim(paste0(root,"/PRISM_repurposing_alpelisib_copanlisib.txt"))
lineage = fread(paste0(root, "/sample_info.csv"))
```

Calculate the correlation between ΔCDS2 lethality and ΔCDIPT lethality, ΔPIK3CA lethality, PIK3CA inhibitor lethality (Alpelisib) or pan-PI3K inhibitor lethality (Copanlisib) in solid cancer cell lines from DepMap. Use cell lines with available RNA to allow follow-up analyses with gene expression data.
```{r}
rownames(CERES_raw) = CERES_raw$DepMap_ID
rownames(drugs) = drugs$X

genes = grep("CDS2|CDIPT|PIK3CA", colnames(CERES_raw))
lethality = CERES_raw[rownames(CERES_raw) %in% RNA$V1, genes]
lethality = cbind(lethality, drugs[match(rownames(lethality), rownames(drugs)),])

colnames(lethality) = c("CDIPT", "CDS2", "PIK3CA", "depmap_id", "copanlisib", "alpelisib")

# remove blood lineage cell lines as we are using only solid tumors
blood = lineage[lineage=="blood", DepMap_ID]
lethality = lethality[!rownames(lethality) %in% blood,]
```

Do all the correlation tests and make a data.table from the p-values and correlations
```{r}
cdipt = cor.test(lethality[, "CDS2"], lethality[, "CDIPT"], method = "pearson", alternative = "two.sided")
pik3ca = cor.test(lethality[, "CDS2"], lethality[, "PIK3CA"], method = "pearson", alternative = "two.sided")
copanlisib = cor.test(lethality[!is.na(lethality$copanlisib), "CDS2"], lethality[!is.na(lethality$copanlisib), "copanlisib"], method = "pearson", alternative = "two.sided")
alpelisib = cor.test(lethality[!is.na(lethality$alpelisib), "CDS2"], lethality[!is.na(lethality$alpelisib), "alpelisib"], method = "pearson", alternative = "two.sided")

plot = data.table(comparison = c("cdipt", "pik3ca", "copanlisib", "alpelisib"),
                  pval = c(cdipt$p.value, pik3ca$p.value, copanlisib$p.value, alpelisib$p.value),
                  r = c(cdipt$estimate, pik3ca$estimate, copanlisib$estimate, alpelisib$estimate))
plot[, padj := p.adjust(pval, method = "bonferroni")]
```

